### 2021-08-05 paper/paper.tex
make some minor tweaks (references, when abbreviations are introduced etc.)

@@

\todo{TODO: Add MHD results?}

@@

Many scientists implementing numerical methods often have no real experience or
education in software development. Julia and its ecosystem support these
researchers by making it easy to set up unit and regression tests, since a
testing framework is included in the standard library and default package
generation setups prepare the file structure accordingly. This allows test
driven development and continuous integration (CI), which makes code restructuring
and optimization straightforward.

@@

Similarly, there is no need to compile HDF5 on a user system due to the
availability of HDF5.jl\footnote{\url{https://github.com/JuliaIO/HDF5.jl}}.
There is still work to be done how to improve the experience in combination
with MPI on high performance computing (HPC) clusters, but some promising
approaches exist \cite{byrne2021mpi}.

@@

Due to Julia's package manager, the cost of using external dependencies is low
enough to not be a problem. Traditionally, many scientific simulation codes tend
to reduce the number of external dependencies as much as possible due to the
complexity of handling different versions and making everything work together.
Such an approach typically leads to significant code duplication. For example, many
CFD codes implement their own time integration methods, arguing that the spatial
part is much more complex. While this is often true, it is still more efficient
in terms of human workload to be able to reuse existing implementations. This
allows experts on time integration methods to develop specialized algorithms and
implement them in open source software while practitioners or researchers focusing
on spatial semidiscretizations ca benefit immediately by changing only one line
of code. For example, optimized time integration methods were developed in
\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. These methods
can be used with \trixi by changing a single line of code. In contrast, researchers
working without external dependencies would need to spend a day or two to digest
details before even beginning implementation. This additional overhead in developer
time makes it less likely that researchers would implement these methods in their own 
codes and benefit from recent algorithmic developments.

@@

Due to Julia's high-level programming approach, dynamic typing, and multiple
dispatch, it is easy to combine existing functionality efficiently. For example,
the single-physics solvers for hyperbolic PDEs of \trixi were extended to
a multi-physics setup for the compressible Euler equations with self-gravity
with roughly 350 lines of code \cite{schlottkelakemper2021purely}.
In addition, \trixi can be extended from the outside without modifying the main
source code, which makes it easy to set up new simulation approaches and analyze
existing ones, \eg by fluctuation simulations \cite{ranocha2021preventing}.

@@

Based on our experience, we consider Julia to be suitable for simulation-focused
scientific computing, in particular for hyperbolic PDEs, computational fluid
dynamics, and related problems --- at least on the scale of shared memory
parallelism. The scalability of high-order methods for hyperbolic PDEs written
in Julia to high performance computing applications still needs to be demonstrated.
However, we do not expect this to remain an unsolvable problem since Julia is not 
generically slower than traditional compiled HPC programming languages. Nevertheless, 
it appears to be more complicated to scale Julia to HPC while still keeping the
code as simple, clean, and flexible as it is now.

>>>>

\todo{TODO: Add MHD results}

@@

@@

Many scientists implementing numerical methods often have no real experience or
education in software development. Julia and its ecosystem support these
researchers by making it easy to set up unit and regression tests, since a
testing framework is included in the standard library and default package
generation setups prepare the file structure accordingly. This allows test-driven
development and continuous integration (CI), which makes code restructuring
and optimization straightforward.

@@

Similarly, there is no need to compile HDF5 on a user system due to the
availability of HDF5.jl\footnote{\url{https://github.com/JuliaIO/HDF5.jl}}.
There is still work to be done how to improve the experience in combination
with MPI on HPC clusters, but some promising
approaches exist \cite{byrne2021mpi}.

@@

Due to Julia's package manager, the cost of using external dependencies is low
enough to not be a problem. Traditionally, many scientific simulation codes tend
to reduce the number of external dependencies as much as possible due to the
complexity of handling different versions and making everything work together.
Such an approach typically leads to significant code duplication. For example, many
CFD codes implement their own time integration methods, arguing that the spatial
part is much more complex. While this is often true, it is still more efficient
in terms of human workload to be able to reuse existing implementations. This
allows experts on time integration methods to develop specialized algorithms and
implement them in open source software while practitioners or researchers focusing
on spatial semidiscretizations can benefit immediately by changing only one line
of code. For example, optimized time integration methods were developed in
\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. These methods
can be used with \trixi by changing a single line of code. In contrast, researchers
working without external dependencies would need to spend a day or two to digest
details before even beginning implementation. This additional overhead in developer
time makes it less likely that researchers would implement these methods in their own 
codes and benefit from recent algorithmic developments.

@@

Due to Julia's high-level programming approach, dynamic typing, and multiple
dispatch, it is easy to combine existing functionality efficiently. For example,
the single-physics solvers for hyperbolic PDEs of \trixi were extended to
a multi-physics setup for the compressible Euler equations with self-gravity
with roughly 350 lines of code \cite{schlottkelakemper2021purely}.
In addition, \trixi can be extended from the outside without modifying the main
source code, which makes it easy to set up new simulation approaches and analyze
existing ones, \eg, by fluctuation simulations \cite{ranocha2021preventing}.

@@

Based on our experience, we consider Julia to be suitable for simulation-focused
scientific computing, in particular for hyperbolic PDEs, computational fluid
dynamics, and related problems --- at least on the scale of shared memory
parallelism. The scalability of high-order methods for hyperbolic PDEs written
in Julia to high-performance computing applications still needs to be demonstrated.
However, we do not expect this to remain an unsolvable problem since Julia is not 
generically slower than traditional compiled HPC programming languages. Nevertheless, 
it appears to be more complicated to scale Julia to HPC while still keeping the
code as simple, clean, and flexible as it is now.
